---
title: "p8105_hw2_dr3168"
author: "Daniela Rochez"
date: "10/8/2021"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---


```{r}
library(tidyverse)
library(readxl)
```

#Loading Mr.Trash Wheel Data and Cleaning it
```{r}
trash_wheel_data=read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range =cell_cols("A:N"))
trash_wheel_data= janitor::clean_names(trash_wheel_data)
trash_wheel_data
```
#Omit some missing data
```{r}
trash_wheel_data_missing_values=drop_na(trash_wheel_data, dumpster)
trash_wheel_data_missing_values
```
#Rounding sports balls and finding the median
```{r}
trash_wheel_sports_balls=mutate(trash_wheel_data_missing_values,sports_balls= round(sports_balls))
trash_wheel_sports_balls= as.numeric(trash_wheel_sports_balls %>%pull(sports_balls))
median(trash_wheel_sports_balls)
```

#Reading Precipitation in 2019 and cleaning it
```{r}
Precip_2019= read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "2019 Precipitation", range="A2:B14")
Precip_2019=janitor::clean_names(Precip_2019)
Precip_2019
```

#Reading Precipitation in 2018 and cleaning it
```{r}
Precip_2018=read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range="A2:B14")
Precip_2018=janitor::clean_names(Precip_2018)
Precip_2018
```
#Ommitting missing data
```{r}
Precip_2019_Missing= drop_na(Precip_2019)
Precip_2019_Missing
Precip_2018_Missing= drop_na(Precip_2018)
Precip_2018_Missing
```

#Adding a variable
```{r}
Precip_2019_year=mutate(Precip_2019_Missing, year=2019)
Precip_2019_year
Precip_2018_year=mutate(Precip_2018_Missing, year=2018)
Precip_2018_year
```

#Combining both data sets
```{r}
Precipitation_2019_2018=full_join(Precip_2019_year, Precip_2018_year)
Precipitation_2019_2018
```

#Using month.name
```{r}
mutate(Precipitation_2019_2018, month=month.name[month])
```


#The number of observations in the precipation data set is 24 and one of they key variables is the total precipiation per month. The total precipiation data for the year 2018 is 70.33 inches. In the Mr. Trash Wheel data, the the number of obervations were 406 with 14 variables. These variables ranged from weight of trash in tons to volume of the trash to the amount of sport balls. The median number of sports balls is 8.


#Uploading and cleaning pols data
```{r}
pols_month_df= 
  read_csv("fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month= as.numeric(month)) %>% 
  mutate(year= as.numeric(year)) %>% 
  mutate(month=month.name[month]) %>% 
  pivot_longer(
    cols = starts_with("prez"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "gop_dem") %>% 
  select(-day)

pols_month_df
view(pols_month_df)
```
#Uploading and cleaning snp data

```{r}
snp_df= 
  read_csv("fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date=lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>%
  mutate(year=as.numeric(year)) %>%
  mutate(year=ifelse(year>2021,year-100, year)) %>% 
  mutate(month= as.numeric(month)) %>% 
  mutate(month=month.name[month]) %>% 
  arrange(year,month) %>% 
  relocate(year,month) %>% 
  select(-day)

  
snp_df
view(snp_df)
```

#Uploading and cleaning unemployment data
```{r}
unemployment_df= 
  read_csv("fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec,
    names_to="month",
    values_to="rate") %>% 
  mutate(month= recode(month,`jan`=1, `feb`=2,`mar`=3,`apr`=4,
                       `may`=5,`jun`=6,`jul`=7,`aug`=8,
                       `sep`=9,`oct`=10,`nov`=11,`dec`=12 )) %>% 
  mutate(month=month.name[month])
unemployment_df
view(unemployment_df)
```

#Joining data of snp and pols
```{r}
snp_pols_df=
  left_join(snp_df,pols_month_df, by=c("year", "month"))
snp_pols_df
```

#Joining unemployment data to snp_pols_df
```{r}
all_data=
  left_join(unemployment_df, snp_pols_df, by=c("year", "month"))
all_data
view(all_data)
```

#In the Pols_Month dataset, there were 1,634 observations and 10 variables. A key variable in this set was month and year. In addition to this, president was a key variable and the data shows weather he was republican or democrat based on the year. The range of years for this set was 1947 to 2015. For the Snp dataset, there were 787 observations and 3 variables. The key variables were month,year, and the closing values. The range of years were from 1950 to 2015.The last set, which is the unemployment data, had 816 observations and 3 variables. They key varaibles were year, month and precentage of unemployment. The year range is from 1948  to 2015. When the data were joint, the outcome had 1,602 observations and 12 variables. The year ranged from 1948 to 2015 and contained all the variables from the previous three data. This data set was joined based on two coloumns: year and month. There are a lot of NAs.
